---
title: "World Happiness Analysis"
author: "Tae Yi Ham, Aneesh Reddy, Sean Kim"
date: "2023-12-06"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(countrycode)
library(ggplot2)
library(corrplot)
library(dplyr)
```

## Project Overview

This project uses the 2019 World Happiness Report dataset to examine associations between national happiness scores and GDP per capita, social support, healthy life expectancy, generosity, and geographic region. Because the data are observational and cross-sectional, the analysis describes associations rather than causal effects.

## Load and Prepare the Data

```{r}
happiness <- read_csv("data/2019.csv", show_col_types = FALSE)

names(happiness) <- c(
  "HappinessRank", "Country", "Happiness", "GDP", "SocialSupport",
  "HealthyLife", "Freedom", "Generosity", "Corruption"
)

happiness <- happiness %>%
  mutate(
    Territory = countrycode(Country, origin = "country.name", destination = "region"),
    ISO2C = countrycode(Country, origin = "country.name", destination = "iso2c")
  )

str(happiness)
```

## Pearson Correlation Analysis

```{r}
numeric_data <- happiness %>%
  select(Happiness, GDP, SocialSupport, HealthyLife, Freedom, Generosity, Corruption)

correlation <- cor(numeric_data, use = "complete.obs")
corrplot(correlation, type = "upper", method = "color")

cor(happiness$Happiness, happiness$GDP, use = "complete.obs")
cor(happiness$Happiness, happiness$HealthyLife, use = "complete.obs")
cor(happiness$Happiness, happiness$SocialSupport, use = "complete.obs")
cor(happiness$Happiness, happiness$Generosity, use = "complete.obs")
```

GDP, healthy life expectancy, and social support show the strongest positive bivariate correlations with happiness among the variables emphasized in the original project. Generosity has a much weaker bivariate correlation.

## Multiple Linear Regression

```{r}
model <- lm(
  Happiness ~ SocialSupport + GDP + HealthyLife + Generosity,
  data = happiness
)
summary(model)
```

The model explains approximately 74.2% of the observed variation in happiness scores (R² ≈ 0.742). These coefficients describe statistical associations within this dataset and should not be interpreted as causal effects.

## Scatterplots

```{r}
plot_relationship <- function(data, x, x_label) {
  ggplot(data, aes(x = .data[[x]], y = Happiness)) +
    geom_point(alpha = 0.65) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = paste("Happiness vs.", x_label), x = x_label, y = "Happiness Score") +
    theme_minimal()
}

plot_relationship(happiness, "GDP", "GDP per Capita")
plot_relationship(happiness, "HealthyLife", "Healthy Life Expectancy")
plot_relationship(happiness, "SocialSupport", "Social Support")
plot_relationship(happiness, "Generosity", "Generosity")
```

## Regional Comparison

```{r}
ggplot(happiness %>% filter(!is.na(Territory)), aes(x = Happiness, y = Territory)) +
  geom_boxplot() +
  labs(
    title = "Happiness Score by Region",
    x = "Happiness Score",
    y = "Region"
  ) +
  theme_minimal()
```

The regional distributions differ, but sample sizes are uneven across regions. In particular, regions represented by only a small number of countries should be interpreted cautiously.

## Limitations

This project uses a single year of observational data. Happiness is subjective, regional sample sizes are uneven, and the analysis cannot establish causal relationships. The 2019 data also predate the COVID-19 pandemic.
